---
title: "Falcon Airlines"
author: "Duchoma Agoons"
date: "22/03/2023"
output:
  pdf_document: default
  html_document: default
---

#Data Importation
```{r}
getwd()
library(readr)
Falcon_Data=read.csv("Falcon Customers Data.csv")
sum(is.na(Falcon_Data)) # our dataset contains 23,888 missing enteries. Customers did not provide answers. 
```


#Exploratory steps and data cleaning
```{r}
head(Falcon_Data)
tail(Falcon_Data)
names(Falcon_Data) #We can actually look at which columns are numeric and which are categorical. see next code
colnames(Falcon_Data[,sapply(Falcon_Data,is.numeric)])
colnames(Falcon_Data[,sapply(Falcon_Data,is.factor)])


# Rename Variables
colnames(Falcon_Data)[12]<-"Departure_or_Arrival_Time_Convenience"
colnames(Falcon_Data)[8]<-"Departure_Delay_in_Mins"
colnames(Falcon_Data)[9]<-"Arrival_Delay_in_Mins"
colnames(Falcon_Data)[15]<-"Inflight_Wifi_Service"
names(Falcon_Data)

# ************************** ALTERNATIVE CODE TO CLEANING VARIABLE NAMES ************************************
colnames(Falcon_Data)=make.names(colnames(Falcon_Data))
?make.names

# Rename levels of variables CustomerType, TypeTravel based on the summary function's output. This code 
# replaces an empty space space to "Others". It is being done in two steps. 
summary(Falcon_Data) 
Falcon_Data$CustomerType[Falcon_Data$CustomerType == ""] <- NA
Falcon_Data$CustomerType = as.factor(replace(as.character(Falcon_Data$CustomerType), is.na(Falcon_Data$CustomerType),"Others"))

Falcon_Data$TypeTravel[Falcon_Data$TypeTravel == ""] <- NA
Falcon_Data$ TypeTravel = as.factor(replace(as.character(Falcon_Data$ TypeTravel), is.na(Falcon_Data$ TypeTravel),"Others"))

#### Observation: Variables were cleaned correctly. 



anyNA(Falcon_Data)
attach(Falcon_Data)


# Overview of data
str(Falcon_Data)
library(tidyverse)
glimpse(Falcon_Data)

library(Hmisc)
describe(Falcon_Data)
# Describes data very well. The output from here will be useful in interpreting "plot_histogram(Falcon_Data)" output.

``` 



# Create bins of Age variable  - Manual splitting using the cut function
## For this project, we shall not label the age bins for easy visual comprehension. Labelling the age groups
## may confuse stakeholders. Nevertheless, the definition of the Age groups are as follows:
## This is an engineering exercise. You need to try splitting the variable into various sizes. Some granular details will be revealed.

### Observation
# 1) This bins splitting resulted in telling us one more granular detail in comparison to when we let R split the data for us.
#   Departure_or_Arrival_time_convenience was highly rated by "Kids", which is only but normal. They have virtually no time constraints.

```{r}
Age.Bins=cut(Age,c(7,12,19,29,49,60,69,85),include.lowest=TRUE) 
table(Age.Bins) 
plot(Age.Bins,col="light green") 

# Now, we add the Age.Bins to our Dataset 
Falcon_Data$Age.Bins=Age.Bins 
attach(Falcon_Data) 

# Segmentation of Data by Age.Bins 
library(DataExplorer) 

## We use the "Age.bins" variables to extract more information from the data.
## Deeper understanding of customer Age groups

table(Age.Bins,Satisfaction) 
prop.table(table(Age.Bins,Satisfaction)) 
plot(Age.Bins,Satisfaction,col=c("light green", "light blue")) 
### Observation:  Plot this using Tableau. 


table(Age.Bins,CustomerType)
prop.table(table(Age.Bins,CustomerType))
plot(Age.Bins,CustomerType,col=c("light green","light blue","violet"))
### Observation: 


table(Age.Bins,TypeTravel)
prop.table(table(Age.Bins,TypeTravel))
plot(Age.Bins,TypeTravel,col=c("light green","light blue","violet"))
### Observation: 

table(Age.Bins,Class)
prop.table(table(Age.Bins,Class))
plot(Age.Bins,Class,col=c("light green","light blue","violet"))
### Observation: 


table(Age.Bins,Gender)
prop.table(table(Age.Bins,Gender))
plot(Age.Bins,Gender,col=c("light green","light blue")) # Approximately the same proportion os men and women are in all buckets.
### Observation:  No significant differentiation. 

plot(Age.Bins,Flight_Distance,col=c("light blue","orange","yellow","green","pink","violet","red"))
```

## NOTE
# (i) We need to do line charts
# (ii) We need to show number of observations in each bucket of our bar graphs. This will shine more light on the percentage
#    representation we we use to describe our bar diagrams.
# (iii) Show the percentage of each bars in our bar graph


# Bar Charts for Categorical variables
```{r}
library(ggplot2)
library(dplyr)

#******************************************** Customer Type ****************************************************************#

CustomerType_count=Falcon_Data %>% count(CustomerType) # This creates a frequency table for the CustomerType variable.

# This code helps represent the counts on each bucket
ggplot(CustomerType_count, aes(x= reorder(CustomerType,-n), y=n)) + 
  geom_bar(stat="identity",fill="light green", color="black") + 
  geom_text(aes(label=n), vjust=-0.25) + 
  labs(x="CustomerType", y="Count", title="Customer Type Distribution")

# Represent percentage rather than count on buckets
## begin by creating a new data frame with percentage equivalents
Percentage_CustomerType = Falcon_Data %>% 
  count(CustomerType) %>% 
          mutate(per = n / sum(n),
                 per_label = paste0(round(per*100), "%"))

ggplot(Percentage_CustomerType, aes(x = reorder(CustomerType, -per), y = per)) + 
  geom_bar(stat = "identity", fill = "light green", color = "black") +
    geom_text(aes(label = per_label), vjust = -0.25) +
    labs(x = "CustomerType", y = "Count", title = "Customer Type Distribution")+ 
    scale_y_continuous(labels = scales::percent) +
    theme_bw()
#************************************************************************************************************************#


#******************************************** Gender  ****************************************************************#

## This code prints the number of observations on the chart bars
Gender_count=Falcon_Data %>% count(Gender) 

ggplot(Gender_count, aes(x= reorder(Gender,-n), y=n)) + 
  geom_bar(stat="identity",fill="light green", color="black") + 
  geom_text(aes(label=n), vjust=-0.25) + 
  labs(x="Gender", y="Count", title="Gender Distribution")


# This code prints the percentage representation of data on the chart bar
Percentage_Gender = Falcon_Data %>% 
  count(Gender) %>% 
          mutate(per = n / sum(n),
                 per_label = paste0(round(per*100), "%"))
 
ggplot(Percentage_Gender, aes(x = reorder(Gender, -per), y = per)) + 
  geom_bar(stat = "identity", fill = "light green", color = "black") +
    geom_text(aes(label = per_label), vjust = -0.25) +
    labs(x = "Gender", y = "Count", title = "Gender Distribution")+ 
    scale_y_continuous(labels = scales::percent) +
    theme_bw()
#************************************************************************************************************************#


##******************************************** Age.Bins ****************************************************************#
## This code prints the number of observations on the chart bars
Age.Bins_count=Falcon_Data %>% count(Age.Bins) 

ggplot(Age.Bins_count, aes(x= reorder(Age.Bins,-n), y=n)) + 
  geom_bar(stat="identity",fill="light green", color="black") + 
  geom_text(aes(label=n), vjust=-0.25) + 
  labs(x="Age.Bins", y="Count", title="Age.Bins Distribution")


# This code prints the percentage representation of data on the chart bar
Percentage_Age.Bins = Falcon_Data %>% 
  count(Age.Bins) %>% 
          mutate(per = n / sum(n),
                 per_label = paste0(round(per*100), "%"))
 
ggplot(Percentage_Age.Bins, aes(x = reorder(Age.Bins, -per), y = per)) + 
  geom_bar(stat = "identity", fill = "light green", color = "black") +
    geom_text(aes(label = per_label), vjust = -0.25) +
    labs(x = "Age.Bins", y = "Count", title = "Age.Bins Distribution")+ 
    scale_y_continuous(labels = scales::percent) +
    theme_bw()
#************************************************************************************************************************#


#******************************************** Type Travel ****************************************************************#
## This code prints the number of observations on the chart bars
TypeTravel_count=Falcon_Data %>% count(TypeTravel) 

ggplot(TypeTravel_count, aes(x= reorder(TypeTravel,-n), y=n)) + 
  geom_bar(stat="identity",fill="light green", color="black") + 
  geom_text(aes(label=n), vjust=-0.25) + 
  labs(x=" TypeTravel ", y="Count", title=" TypeTravel Distribution")


# This code prints the percentage representation of data on the chart bar
Percentage_TypeTravel = Falcon_Data %>% 
  count(TypeTravel) %>% 
          mutate(per = n / sum(n),
                 per_label = paste0(round(per*100), "%"))
  
ggplot(Percentage_TypeTravel, aes(x = reorder(TypeTravel, -per), y = per)) + 
  geom_bar(stat = "identity", fill = "light green", color = "black") +
    geom_text(aes(label = per_label), vjust = -0.25) +
    labs(x = " TypeTravel ", y = "Count", title = " TypeTravel Distribution")+ 
    scale_y_continuous(labels = scales::percent) +
    theme_bw()
#************************************************************************************************************************#


#******************************************** Class ****************************************************************#
## This code prints the number of observations on the chart bars
Class_count=Falcon_Data %>% count(Class) 

ggplot(Class_count, aes(x= reorder(Class,-n), y=n)) + 
  geom_bar(stat="identity",fill="light green", color="black") + 
  geom_text(aes(label=n), vjust=-0.25) + 
  labs(x="Class", y="Count", title=" Class Distribution")


# This code prints the percentage representation of data on the chart bar
Percentage_Class = Falcon_Data %>% 
  count(Class) %>% 
          mutate(per = n / sum(n),
                 per_label = paste0(round(per*100), "%"))
    
ggplot(Percentage_Class, aes(x = reorder(Class, -per), y = per)) + 
  geom_bar(stat = "identity", fill = "light green", color = "black") +
    geom_text(aes(label = per_label), vjust = -0.25) +
    labs(x = " Class ", y = "Count", title = " Class Distribution")+ 
    scale_y_continuous(labels = scales::percent) +
    theme_bw()
#************************************************************************************************************************#


#******************************************** Satisfaction ****************************************************************#
## This code prints the number of observations on the chart bars
Satisfaction_count=Falcon_Data %>% count(Satisfaction) 

ggplot(Satisfaction_count, aes(x= reorder(Satisfaction,-n), y=n)) + 
  geom_bar(stat="identity",fill="light green", color="black") + 
  geom_text(aes(label=n), vjust=-0.25) + 
  labs(x="Class", y="Count", title="Satisfaction Distribution")


# This code prints the percentage representation of data on the chart bar
Percentage_Satisfaction = Falcon_Data %>% 
  count(Satisfaction) %>% 
          mutate(per = n / sum(n),
                 per_label = paste0(round(per*100), "%"))
    
ggplot(Percentage_Satisfaction, aes(x = reorder(Satisfaction, -per), y = per)) + 
  geom_bar(stat = "identity", fill = "light green", color = "black") +
    geom_text(aes(label = per_label), vjust = -0.25) +
    labs(x = " Satisfaction ", y = "Count", title = "Satisfaction Distribution")+ 
    scale_y_continuous(labels = scales::percent) +
    theme_bw() 
#************************************************************************************************************************#

## The above visuals are a better representation of this joint bar plot
plot_bar(Falcon_Data) 

```



# EDA 
## Challenge: This section can be improved if Stacked bar charts with percentages can be plotted; Tableau was used for this section. 
### W plotted this graphs using Tableau
```{r}

summary(Falcon_Data)

# Customer profiling


# Customer type vs Gender
table(Gender,CustomerType)
prop.table(table(Gender,CustomerType))
plot(Gender,CustomerType,col=c("light green", "light blue","violet")) #Differentiation

# TypeTravel vs Departure_Delay_in_Mins
plot(TypeTravel,Departure_Delay_in_Mins)
### Observation:  no differentiation  but for number of outliers

# Gender vs Class
table(Gender,Class)
prop.table(table(Gender,Class))
plot(Gender, Class, col=c("light green", "light blue","violet"))
### Observation: There is no diffeentiation in the preference of Males or females and the class they like flighing. 

# Gender vs Satisfaction 
table(Gender,Satisfaction)
prop.table(table(Satisfaction))
plot(Gender,Satisfaction,col=c("light green", "light blue")) 
### Obervation: More females are satisfied than Males. After regression, we will see what parameter contributes to satisfaction.


# TypeTravel vs Gender
table(TypeTravel,Gender)
prop.table(table(TypeTravel,Gender))
plot(TypeTravel,Gender, col=c("light green", "light blue","violet")) 
### Observation: No differentiation 


# CustomerType vs Satisfaction
table(CustomerType,Satisfaction)
prop.table(table(CustomerType,Satisfaction))
plot(CustomerType,Satisfaction,col=c("light green", "light blue"))
### Observation: 45% of loyal customers are satisfied while 28% disatisfied. 39% of disloyal customers are satisfied while 12.5% are dissatisfied.
### Observation:  For other customers, overall, they are satisfied but the satisfaction - dissatisfaction margin is too narrow. 


# CustomerType vs Class
table(CustomerType,Class)
prop.table(table(CustomerType,Class))
plot(CustomerType,Class,col=c("light green", "light blue","violet"))
### Observation:  36.8% of loyal customers fly business, 30.8% of them fly Eco, and 5.9% of them fly Eco Plus.
### Observation: 6.4% of disloyal customers fly business, the majority of disloyal customers fly Eco class (9.7%), and 0.6% of them fly Eco Plus.


# TypeTravel vs Satisfaction
table(TypeTravel,Satisfaction)
prop.table(table(TypeTravel,Satisfaction))
plot(TypeTravel,Satisfaction,col=c("light green", "light blue"))
### Observation:  The majority of business class travelers are satisied. 
### 36% of business class travelers are satisfied while 25.8% of them are dissatisfied.
### Observation:  The majority of personal type travelers are not satisfied (15%) while 13% are satisfied.


# TypeTravel vs Class
table(Class,TypeTravel)
prop.table(table(Class,TypeTravel))
plot(Class,TypeTravel,col=c("light green", "light blue","violet"))
### Observation: Most people who travel for business use Business class. Most people who travel for private reasons use Eco class.
### 41% of those who travel for business purposes use business class, 4.8% of people who travel for other reasons use business class and 
## 1.9% of those who travel for personal reasons use business class. 
### Observation: 17.5% of people who travel for  for business purposes use Eco class and 22.8% of those who travel for private reasons use Eco.
### Observation: 6624 clients travel by Eco Plus class (7.3% of customers).


# Class vs Satisfaction 
table(Satisfaction, Class)
prop.table(table(Satisfaction, Class))
plot(Satisfaction, Class,col=c("light green", "light blue","violet"))
### Observation:  The majority of business class travelers are satisfied (34%), while 13.9% are dissatisfied.
### Observation: 27.2% of Eco class passengers are not satified while 17.6% are.
### Observation: The majority of Eco Plus (4.1%) passengers are dissatisfied while 3.1% of passengers are. 


# Class vs Flight_Distance
plot(Class,Flight_Distance,col=c("light green","light blue","violet"))
### Observation: Business class passengers travel on average higher miles than other passengers in the Eco and Eco Plus class.
```


#EDA and segmentation of our data by the flight data variables.
# Don't worry about the segmentation of Age for the other algorithms.

```{r}

library(DataExplorer)
plot_str(Falcon_Data)
plot_intro(Falcon_Data) # Our data contains missing observations. 
plot_missing(Falcon_Data) #Data contains NA's 
profile_missing(Falcon_Data) # Shows us the which variables contain missing values
plot_histogram(Falcon_Data)
plot_correlation(Falcon_Data,type='continuous') # Our correlation doesn't show any relationship between variables which is actionable.
# We change the dependent variable, Satisfaction, to a numeric and then re-run the correlation. It yieled the same results. 


plot_bar(Falcon_Data) # The majority of clients are in the age group "29-49"


plot_boxplot(Falcon_Data,by="Satisfaction")
plot_boxplot(Falcon_Data,by="Age")
plot_boxplot(Falcon_Data,by="Gender")
plot_boxplot(Falcon_Data,by="CustomerType")
plot_boxplot(Falcon_Data,by="Age.Bins")
plot_boxplot(Falcon_Data,by="Class")
plot_boxplot(Falcon_Data,by="Flight_Distance")
plot_boxplot(Falcon_Data,by="Departure_Delay_in_Mins")
plot_boxplot(Falcon_Data,by="Arrival_Delay_in_Mins")
plot_boxplot(Falcon_Data,by="Departure_or_Arrival_Time_Convenience")
plot(Departure_or_Arrival_Time_Convenience,Age.Bins)
```



#Exploring the survey data
```{r}
plot_boxplot(Falcon_Data,by="Seat_comfort")
plot_boxplot(Falcon_Data,by="Departure_or_Arrival_Time_Convenience")
plot_boxplot(Falcon_Data,by="Food_drink")
plot_boxplot(Falcon_Data,by="Gate_location")
plot_boxplot(Falcon_Data,by="Inflight_Wifi_Service")#  The majority of low ratings are given by customers in the age group of 20 to 30 years old. 

plot_boxplot(Falcon_Data,by="Inflight_entertainment") # The majority of low ratings are given by customers in the age group of 20 to 30 years old. 
plot_boxplot(Falcon_Data,by="Online_support")
plot_boxplot(Falcon_Data,by="Ease_of_Onlinebooking")
plot_boxplot(Falcon_Data,by="Onboard_service")
plot_boxplot(Falcon_Data,by="Leg_room_service")
plot_boxplot(Falcon_Data,by="Baggage_handling")
plot_boxplot(Falcon_Data,by="Checkin_service")
plot_boxplot(Falcon_Data,by="Cleanliness")
plot_boxplot(Falcon_Data,by="Online_boarding")
```
# Observation: The majority of low ratings are given by customers in the age group of 20 to 30 years old. 


## The DataExplorer library function already returned this graphic as a lump and we have commennted on it already
```{r}
str(Falcon_Data)
hist(Age)
hist(Flight_Distance)
hist(Departure_Delay_in_Mins)
hist(Arrival_Delay_in_Mins)
hist(Seat_comfort)
hist(Departure_or_Arrival_Time_Convenience)
hist(Food_drink)
hist(Gate_location)
hist(Inflight_Wifi_Service)
hist(Inflight_entertainment)
hist(Online_support)
hist(Ease_of_Onlinebooking)
hist(Onboard_service)
hist(Leg_room_service)
hist(Baggage_handling)
hist(Checkin_service)
hist(Cleanliness)
hist(Online_boarding)

```
# Observation:  These pattern recognition diagrams above help us to understand the iindividual parameters of both the flight and survey data. This 
# gives us an idea of where the company stands. 



### Chisuared test to investigate the relationship between categorical data
# Relationship between Categorical data
# Ho: The variables are independent.
# Ha: There is an association between the variables.
# Given alpha = 0.05 as our level of significance, and given the p-values which are all less than 0.05, we 
# overwhelmingly fall to accept the null hypothesis that the variables are independent and thus, conclude that, there is 
# an association or a relationship with the variables tested.
## The survey data parameters all have an association or affect the value customers derive from Falcon airlines company.
```{r}
library(DataExplorer)

chisq.test(Seat_comfort,Leg_room_service)
chisq.test(CustomerType,TypeTravel)
chisq.test(Satisfaction,Seat_comfort)
chisq.test(Satisfaction,TypeTravel)
chisq.test(Satisfaction,Class)
chisq.test(Satisfaction,Departure_or_Arrival_Time_Convenience)
chisq.test(Satisfaction,Food_drink)
chisq.test(Satisfaction,Inflight_Wifi_Service)


# Chi-squared approximation may be incorrect but however, all variables proved to have a relationship.
chisq.test(Satisfaction,Gate_location)
chisq.test(Cleanliness,Food_drink)
chisq.test(Satisfaction,Online_support)
chisq.test(Satisfaction,Online_boarding)
chisq.test(Satisfaction,Onboard_service)
chisq.test(Satisfaction,Checkin_service)
chisq.test(Satisfaction,Cleanliness)
chisq.test(Inflight_Wifi_Service,Inflight_entertainment)
chisq.test(Gate_location,Departure_or_Arrival_Time_Convenience)
chisq.test(Onboard_service,Checkin_service)
chisq.test(Online_support,Online_boarding)
chisq.test(Online_boarding,Ease_of_Onlinebooking)
```
# We have this error notification because some buckers might have small amounts of data but this is not a problem.


# Correlation  and Histogram plots
```{r}
cor.test(Departure_Delay_in_Mins,Arrival_Delay_in_Mins)
scatter.smooth(Departure_Delay_in_Mins,Arrival_Delay_in_Mins) # This plot in contained in the one just below. 

library(car) 
scatterplot(Departure_Delay_in_Mins,Arrival_Delay_in_Mins)


library(psych)
pairs.panels(Falcon_Data[4:9],gap=0,bg=c("green","blue","yellow")[Falcon_Data$Satisfaction],pch=21)
Falcon_Data$Satisfaction=as.numeric(Falcon_Data$Satisfaction)


# Transform "Satisfaction back to factor
Falcon_Data$Satisfaction=as.factor(Falcon_Data$Satisfaction)

```
# Observation: 
# There exsists  a high correlation between Departure_Delay_in_Mins and Arrival_Delay_in_Mins. However, the majority of flights take off and 
# arrive on time.  We need to delete one of the highly correlated variables.



# Outlier  detection and treatment: This code identifies the inter-quartile range for the variables with outliers and caps it by the
# inter-quartile range values. 
```{r}
boxplot(Departure_Delay_in_Mins) # contains extreme values
boxplot(Flight_Distance) # contains extreme values
boxplot(Age)
boxplot(ID)
boxplot(Arrival_Delay_in_Mins) # contains estreme values 
boxplot(Onboard_service) # contains 2 outliers 


# Observation: Five variables contains outliers

#----------------------------------------------------------------------------------------------------
#Identifying the outlier boundaries
boxplot(Departure_Delay_in_Mins)

IQRage=IQR(Falcon_Data$Departure_Delay_in_Mins)
LLage=quantile(Falcon_Data$Departure_Delay_in_Mins,0.25) - 1.5*IQRage
ULage=quantile(Falcon_Data$Departure_Delay_in_Mins,0.75) + 1.5*IQRage
Departure_Out=subset(Falcon_Data,Departure_Delay_in_Mins < LLage | Departure_Delay_in_Mins > ULage)
dim(Departure_Out)
# 12620 outliers detected
# The upper value is 30. we should cap the value by 30
Departure_WOUT=subset(Falcon_Data, Departure_Delay_in_Mins >= LLage & Departure_Delay_in_Mins <= ULage)
max(Departure_WOUT$Departure_Delay_in_Mins) # answer returned "30", so we use it below.
dim(Departure_WOUT)

summary(Departure_Delay_in_Mins)
Falcon_Data$Departure_Delay_in_Mins[Falcon_Data$Departure_Delay_in_Mins>30]=30

#------------------------------------------------------------------------------------------------------

#------------------------------------------------------------------------------------------------------
boxplot(Flight_Distance)
IQRage=IQR(Falcon_Data$Flight_Distance,na.rm=TRUE)
LLage=quantile(Falcon_Data$Flight_Distance,0.25) - 1.5*IQRage
ULage=quantile(Falcon_Data$Flight_Distance,0.75) + 1.5*IQRage
Flight_Distance_Out=subset(Falcon_Data,Flight_Distance < LLage | Flight_Distance > ULage)
dim(Flight_Distance_Out)

# Capping of outliers
Flight_Distance_WOUT=subset(Falcon_Data,Flight_Distance >= LLage & Flight_Distance <= ULage)
dim(Flight_Distance_WOUT)
max(Flight_Distance_WOUT$Flight_Distance) 

summary(Flight_Distance)
Falcon_Data$Flight_Distance[Falcon_Data$Flight_Distance>4315]=4315

# 1800 outliers detected
#-------------------------------------------------------------------------------------------------------

#-------------------------------------------------------------------------------------------------------
boxplot(Arrival_Delay_in_Mins)

IQRage=IQR(Falcon_Data$Arrival_Delay_in_Mins,na.rm=TRUE)
LLage=quantile(Falcon_Data$Arrival_Delay_in_Mins,0.25,na.rm=TRUE) - 1.5*IQRage
ULage=quantile(Falcon_Data$Arrival_Delay_in_Mins,0.75,na.rm=TRUE) + 1.5*IQRage
Arrival_Out=subset(Falcon_Data,Arrival_Delay_in_Mins < LLage | Arrival_Delay_in_Mins > ULage)
dim(Arrival_Delay_in_Mins)
# 1800 outliers detected

# Capping of outliers
Arrival_WOUT=subset(Falcon_Data,Arrival_Delay_in_Mins >= LLage & Arrival_Delay_in_Mins <= ULage)
dim(Arrival_WOUT)
max(Arrival_WOUT$Arrival_Delay_in_Mins)


summary(Arrival_Delay_in_Mins)
Falcon_Data$Arrival_Delay_in_Mins[Falcon_Data$Arrival_Delay_in_Mins>32]=32
#-------------------------------------------------------------------------------------------------------
```

# Outliers were capped correctly. We verified this by re-running the plot_histogram(Falcon_Data) code and the maximum values showed as
# capped in the above codes.





#################### Treatment for missing values using "mice" library.
# Reference https://www.youtube.com/watch?v=An7nPLJ0fsg

# Procedures 
# 1)  This code finds the percentage of missing values in our data and the multiplication by 100 is to put the number 
# of missing data in percentage representation. 

# 2) Next, we apply the above formula  in (1) to the columns of our dataset ( where "2" stands for columns in that formula).
## Observations: 
# Arrival_Delay_in_Mins has 0.31% of missing data
# Food_drink contains 8.99% of missing data
# Onboard_service contains 7.89% of missing data
# Departure_or_Arrival_Time_Convenience contains 9.06% of missing data
 
# 3) Now, lets check the missing data pattern  (missing data pattern = md.pattern) of our dataset. 


# 4) Imputation for missing data. 
# Note: "m = number of imputations. The default is 5 but you can still use m=3
# Note: Always set seed when running mice code.
# The method for dealing with imputing numeric values is pmm=predictive mean matching.  The method used for imputing
# categorical variables is "polyreg=multinomial logistic regression" 
# "imp" stands for imputation. 

# 5) 
#This means that, from the m=3 we chose, we chose to replace our NA's with the 3rd estimated values 
# R suggested to us.

```{r}
# checking number of NA's 
library(mice)


P=function(x){sum(is.na(x))/length(x)*100}
apply(Falcon_Data,2,P)
md.pattern(Falcon_Data) 
impute=mice(Falcon_Data[,-1],m=5,seed=1000) 
complete(impute,3) 

# This is a smart step
Mice.Complete=complete(impute,3)

# Finally, we store the imputed Data as our nnew dataset with which we will continue doing analysis
Falcon_Data=Mice.Complete 
summary(Falcon_Data)
anyNA(Falcon_Data) #verification of imputation process by checking for NA's.

```
#Observation: All the NA's have imputed using the mice method. 


###################### Deletion of multi-collinear variables. The ID column was already deleted above. 
```{r}
Falcon_Data=Falcon_Data[,-8,] 
view(Falcon_Data)
write.csv(Falcon_Data,"Cleaned_FalconData.csv")
```

### Linear regression and multi-collinearity cleaning.
# The important predictors of Satisfaction can be listed in order of importance by looking at the "t-values". Decision trees can also give us this.
# The R-squared value is 48%, telling us that about 48% of Satisfaction is explained by the our data variations.
# Inflight_Entertainment is the most important driver of sales. 
# To improve satisfaction, focus on lower age groups, as indicated by the negative t-value for Age. 
```{r}
Falcon_Data$Satisfaction=as.numeric(Falcon_Data$Satisfaction)

Falcon_Reg=lm(Satisfaction~.,data=Falcon_Data)
summary(Falcon_Reg) # see observation
Falcon_Reg

library(car)
vif(Falcon_Reg)
```
# Since linear regression accepts on numeric variable type, we start by converting the dependent variable to numeric.


#Segmentation of data Procedure 1
```{r}
library(cluster)
library(dplyr)
library(ggplot2)
library(Rtsne)

# Run Gower distance on entire data. This returned an error
gower_dist = daisy(Falcon_Data,metric=c("gower"))

# The above code returned an error message because our data is too large. We split it into 10
# equal groups and ran the code with subsets of the data.
Reduced_FalconData = Falcon_Data[sample(1:nrow(Falcon_Data),nrow(Falcon_Data)/10),] 

gower_dist = daisy(Reduced_FalconData, metric=c("gower"))

# new code 
gower_mat = as.matrix(gower_dist)


# We now search for the best number of clusters in which to cluster our data
sil_width <- c(NA)
for(i in 2:10){
pam_fit<-pam(gower_dist, diss=TRUE, k = i)
  sil_width[i] <- pam_fit$silinfo$avg.width
}

# Plot the Silhouette Width. The highest width will be our chosen number of clusters ( Don't use elbow method here, it doesn't apply)
plot(1:10, sil_width,
     xlab = "Number of clusters",
     ylab = "Silhouette Width")
lines(1:10, sil_width)

#### INTERPRETATION: n = ? has the highest Silhouette width. So, lets pick k = ? as our number of clusters. 


############ SUMMARY OF EACH CLUSTER. We store the results of pam in our dataset we used to run the "gower_dist"

k <- 2
pam_fit <- pam(gower_dist, diss = TRUE, k)
pam_results <- Reduced_FalconData %>%
  mutate(cluster = pam_fit$clustering) %>%
  group_by(cluster) %>%
  do(the_summary = summary(.))
pam_results$the_summary

# Visualize the clusters in a lower dimensional space 
tsne_obj <- Rtsne(gower_dist, is_distance = TRUE)
tsne_data <- tsne_obj$Y %>%
  data.frame() %>%
  setNames(c("X", "Y")) %>%
  mutate(cluster = factor(pam_fit$clustering))
ggplot(aes(x = X, y = Y), data = tsne_data) +
  geom_point(aes(color = cluster))


```


#Segmentation of data Procedure 2 (this is what we will consider for this project)
```{r}
library(cluster)
library(dplyr)
library(ggplot2)
library(Rtsne)

# We begin by splitting our data in the 90 - 10% 
library(caTools)
set.seed(1000)

split=sample.split(Falcon_Data$Satisfaction,SplitRatio=0.90)

Falcon_DontRunClustering=subset(Falcon_Data, split == TRUE)
Falcon_Clustering=subset(Falcon_Data, split == FALSE)

write.csv(Falcon_Clustering,"Falcon_Clustering.csv")


#we can check the ratios of the minority class to majority class
table(Falcon_Clustering$Satisfaction)
print(prop.table(table(Falcon_Clustering$Satisfaction)))
### Observation: The split dataset has the same proportion as he original dataset. 


# Run Gower distance on for 10% of data using the Falcon_Clustering data
gower_dist = daisy(Falcon_Clustering,metric=c("gower"))


# Re-name the gower
gower_mat = as.matrix(gower_dist)


# We now search for the best number of clusters in which to cluster our data
sil_width <- c(NA)
for(i in 2:10){
pam_fit<-pam(gower_dist, diss=TRUE, k = i)
  sil_width[i] <- pam_fit$silinfo$avg.width
}

# Plot the Silhouette Width. The highest width will be our choosen number of clusters ( Don't use elbow method here, it doesn't apply)
plot(1:10, sil_width,
     xlab = "Number of clusters",
     ylab = "Silhouette Width")
lines(1:10, sil_width)

#### INTERPRETATION: n = ? has the highest Silhouette width. So, lets pick k = ? as our number of clusters. 


############ SUMMARY OF EACH CLUSTER. We store the results of pam in our dataset we used to run the "gower_dist"

k <- 8
pam_fit <- pam(gower_dist, diss = TRUE, k)
pam_results <- Falcon_Clustering %>%
  mutate(cluster = pam_fit$clustering) %>%
  group_by(cluster) %>%
  do(the_summary = summary(.))
pam_results$the_summary


# Visualize the clusters in a lower dimensional space 
tsne_obj <- Rtsne(gower_dist, is_distance = TRUE)
tsne_data <- tsne_obj$Y %>%
  data.frame() %>%
  setNames(c("X", "Y")) %>%
  mutate(cluster = factor(pam_fit$clustering))
ggplot(aes(x = X, y = Y), data = tsne_data) +
  geom_point(aes(color = cluster))


#### Make a column for my cluster in the clustering data
Falcon_Clustering$Cluster1=pam_results$cluster


# Create different different data frames from my clusters
df_cluster1=Falcon_Clustering[Falcon_Clustering$Cluster1==1,]
df_cluster2=Falcon_Clustering[Falcon_Clustering$Cluster1==2,]


write.csv(df_cluster1,"Falcon_Cluster1.csv")
write.csv(df_cluster2,"Falcon_Cluster2.csv")

```
######### Observation: 
### The partition of satisfaction variation between Satisfaction and dissatisfaction is preserved. Thus, our random
### sample of 9092 observations is a fair one and we can rely on it and by the central limit theorem, is 
### representative of the 90917 observationsn. However, these stored clusters do not correpsond to that printed
### by  our algorithm.







# Build CART model for segments to see drivers of satisfaction in segments. [NO VALUE ADDED SINCE WE DIDN't HAVE THE RIGHT SEGMENTS]
```{r}
library(rpart)
library(rpart.plot)
set.seed(1000)

tree.Segment=rpart(formula=Satisfaction ~., data=df_cluster1, method = "class", control = rpart.control(minsplit = 4000, minbucket = 1333, cp = 0.0001562))

printcp(tree.Segment)
plotcp(tree.Segment)
rpart.plot(tree.Segment)
```

```{r}
#Get the best cp= tree$cptable[which.min(tree$cptable[,"xerror"]), "CP"]

tree.Segment$cptable
tree.Segment$cptable[,"xerror"]
min(tree.Segment$cptable[,"xerror"])

bestcp.Segment= tree.Segment$cptable[which.min(tree.Segment$cptable[, "xerror"]), "CP"]
bestcp.Segment

print(tree.Segment$cptable)
plot(tree.Segment$cptable)

## Prune the tree based on bestcp ~ ptree=prune(tree,cp=bestcp)
ptree.Segment = prune(tree.Segment, cp=bestcp.Segment)

print(ptree.Segment)
rpart.plot(ptree.Segment, cex = 0.8)



```

## Variable of importance for segment
```{r}
library(caret)

summary(ptree.Segment)
#ptree$variable.importance

df_cart.Segment = data.frame(ptree.Segment$variable.importance)
df_cart.Segment
write.csv(df_cart.Segment,"varimpcart.Segment.csv")
```

Analysis for segment 2 
# Build CART model for segments to see drivers of satisfaction in segments. [NO VALUE ADDED SINCE WE DIDN't HAVE THE RIGHT SEGMENTS]
# Build CART model for segments to see drivers of satisfaction in segments. [NO VALUE ADDED SINCE WE DIDN't HAVE THE RIGHT SEGMENTS]
```{r}
library(rpart)
library(rpart.plot)
set.seed(1000)

tree.Segment00=rpart(formula=Satisfaction ~., data=df_cluster2, method = "class", control = rpart.control(minsplit = 4000, minbucket = 1333, cp = 0.0001562))

printcp(tree.Segment00)
plotcp(tree.Segment00)
rpart.plot(tree.Segment00)
```
