---
title: "Uber Case Study"
author: "Duchoma Agoons"
date: "02/06/2020"
output: html_document
---


### General culture required: Weather and the different parameters used to measure it. 

## This case study talks about Uber pickups in the state of New York for a period of 6 months. Associated variables 
## are given which in our opinion describe what affects those pickups. 
## Dependent variable = pickups 

```{r setup, include=FALSE}
getwd()

# import dataset
Uber=read.csv("Uber Dataset.csv")
attach(Uber)

head(Uber)
tail(Uber)
dim(Uber)
names(Uber)
```

## Our dataset seems to have been well imported. Now, we proceed with some exploratory data analysis to see what the
## data tells us about Uber pickups in the state of New York. Data collected is from 2015-01-01 to 2015-06-30. 

# EDA  
```{r}

library(DataExplorer) 


# Check the structure of the dataset
str(Uber)
plot_str(Uber)

## Our dataset contains 13 variables, 3 categorical and 10 numeric variables

# check missing data -  there are many boroughs which are missing from the data. 10.46% of boroughs that is, 3000 
# missing enteries which is quite significant. We need to take care of this. 
anyNA(Uber)
plot_missing(Uber) 


# High levelvisual of data
plot_intro(Uber)

summary(Uber)
library(Hmisc)
describe(Uber)

# High level visuals 
plot_bar(Uber)

## 

# Histograms - will be looked along sideboxplots 
plot_histogram(Uber)

# Correlation between Manhatan and pickups 
# Correlation between dewp and temp 
plot_correlation(Uber)

plot(borough,pickups)
plot(dewp,temp)
```
### High level overview of data summary

## The wind speed varies from borough to borough. Some boroughs get very low wind speed and others get very high. 

## Road visibility by drivers ranges from 0 to 10 miles across boroughs

## The temperature across boroughs ranges from 2 to 89 degrees F across boroughs. 

## Dew point across boroughs ranges from -16 to 73 degrees F across boroughs.The higher the dew point rises, the greater the 
# amount of moisture in the air. This directly affects how "comfortable" it will feel outside.

## Sea level pressure ranges from  991.4 to 1043.4 across boroughs. low-pressure system moves into an area, it usually leads to cloudiness,
# wind, and precipitation. High-pressure systems usually lead to fair, calm weather. 

## Precipitation is rain, snow, sleet, or hail — any kind of weather condition where something's falling from the sky. In this dataset, it
# 3 variables for precipitation are measured, precipitation for 1h, 6h and 24h and we see how precipitation affects pickups across boroughs.
#  the forms of water particles, whether liquid or solid, that fall from the atmosphere (eg, rain, hail, snow or sleet). 

## Snow depth in inches has a range of 19 inches but differ across boroughs. 

## Within this 6 months period (180 days), 3.9% of the days were public holidays and 96.1% of the days were non public holidays. 

## There are significant amounts of missing boroughs from our data which need to taken care of. 

## There are observations which are very different from all other variables, otherwise known as outliers which need to taken care of.

## The variable Pickup_dt needs to be dealt with separately. We need to break it into its different components - Month, day, weekday and 
# these different components can show us what the state of pickups are within months, days of the month and weekdays. 


## High level summary of data
```{r}

# summary - data contains too many outliers  and this visual is not informative.
plot_boxplot(Uber,by="pickup_dt")

## Manhattan gets the most number of pickups followed by Broklyn and then Queens. On a high level, the weather factors affecting pickups
## seems to follow the same pattern
plot_boxplot(Uber,by="borough") 

# **************************************************************************************************************************************************************
plot_boxplot(Uber,by="pickups")

## Observations
## Pickups  increase as temperatures increase, implies weather gets better. So, when weather is not favourable, people order less.
## We see almost a similar trend with road visibility. Dew point seems to favor pickups, the more the dew point, the more the pickups.
## Positive correlation between pickups and sea level pressure (the higher the pressure, the better the weather)
## The higher the wind speed, the more the pickups. Thus, wind speed doesn't seem to affect pickups. 
# **************************************************************************************************************************************************************

## The lower the average wind speed, the higher the pickups 
plot_boxplot(Uber,by="spd")

## Nothing much to say
plot_boxplot(Uber,by="vsb")

## Nothing much to say on how it relates to pickups 
plot_boxplot(Uber,by="temp")

## Nothing much to say
plot_boxplot(Uber,by="dewp")

## Nothing much to say
plot_boxplot(Uber,by="slp")

## Nothing much to say
plot_boxplot(Uber,by="pcp01")
plot_boxplot(Uber,by="pcp06")
plot_boxplot(Uber,by="pcp24")

## The lower the median snow inches in a borough, the higher the number of pickups
plot_boxplot(Uber,by="sd")
```
## Observations: 
# - Manhattan gets the most number of pickups followed by Broklyn and then Queens. 

## - Pickups  increase as temperatures increase, implies weather gets better. So, when weather is not favourable, people order less.
## - We see almost a similar trend with road visibility. Dew point seems to favor pickups, the more the dew point, the more the pickups.
## - Positive correlation between pickups and sea level pressure (the higher the pressure, the better the weather)
## - The higher the wind speed, the more the pickups. Thus, wind speed doesn't seem to affect pickups negatively. 
## - No clear trend with sea level pressure and pickups. 
## - The lower the median snow inches in a borough, the higher the number of pickups


## How to handle NA's 
```{r}
## This command sums the total number of NA's in any variables in our dataset
sapply(Uber, function(x) sum(is.na(x)))


# This code takes all the NA's and put them in a separate column in our borough variable
Uber$borough = as.factor(replace(as.character(Uber$borough), is.na(Uber$borough),"Unknown B"))
levels(borough)

## This code helps to confirm that the NA's column called "unknown" were successfully added as a 7th borough
table(Uber$borough) 

## This shows us graphically what our borough variable looks like after adding an unknown column to it
plot(aggregate(pickups~borough,data = Uber,sum),type="p")
plot(borough,pickups)

## See  https://www.statmethods.net/graphs/images/lines0.png for graph types other than "p" 

```
## Observations
## All the missing labels of boroughs were grouped and put in particular category. They weren't deleted because they were
## significant. Deleting them meaning deleting observations which in turn might affect out analysis negatively. 


# How to split the Date variable
```{r}
require(pacman)
library(lubridate)
p_load(lubridate)
p_load(dplyr)

# Let us break date into hour, day, month
Uber$start_date = strptime(Uber$pickup_dt,'%Y-%m-%d %H:%M')

# We extract month, day, hour and week day from the newly created date so as to draw more insights at
# a granular level
Uber$start_month = month(Uber$start_date)
Uber$start_day = day(Uber$start_date)
Uber$start_hour = hour(Uber$start_date)
Uber$wday = weekdays(Uber$start_date)
Uber = Uber[,-14]
attach(Uber)

plot_boxplot(Uber,by="start_month")
plot_boxplot(Uber,by="wday")
```
## Observation: 
## - More pickups are experienced on weekends 

## Observation: 
## - 4 new variables have been created: "Start_month", "Start_day", "Start_hour", and "Wday"


## Univariate analysis  for categorical variables and the dependent variable 
```{r}
par(mfrow = c(1,1)); 
text(x = barplot(table(borough),main = "Borough",ylab = "Frequency"), y = 0,  table(borough), cex=1,pos=3); 
text(x = barplot(table(hday),main = "Holiday",ylab = "Frequency"), y = 0,  table(hday), cex=1,pos=3)

```
# Observation: All boroughs rate of pickups is the same with the exception of unspecified boroughs which vary and all these pickups
#       happen 96.1% of the time on non-pyblic holidays


## Univariate analysis for numerical variables and the dependent variable. We plot histograms and boxplots
```{r}
par(mfrow = c(2,2));
text(x= barplot(table(pickups),main = "Pickups",ylab = "Frequency"), y = 0, table(pickups), cex=1,pos=3);
boxplot(pickups, horizontal = TRUE, main = " pickups")

text(x= barplot(table(spd),main = "Wind speed in miles per hour",ylab = "Frequency"), y = 0, table(spd), cex=1,pos=3);
boxplot(spd, horizontal = TRUE, main = "Wind speed in miles per hour")


par(mfrow = c(2,2));
text(x= barplot(table(vsb),main = " Visibility in Miles to nearest tenth",ylab = "Frequency"), y = 0, table(vsb), cex=1,pos=3);
boxplot(vsb, horizontal = TRUE, main = " Visibility in Miles to nearest tenth")

text(x= barplot(table(temp),main = "Temperature in Fahrenheit",ylab = "Frequency"), y = 0, table(temp), cex=1,pos=3);
boxplot(vsb, horizontal = TRUE, main = "Temperature in Fahrenheit");


par(mfrow = c(2,2));
text(x= barplot(table(dewp) ,main = "Dew point in Fahrenheit",ylab = "Frequency"), y = 0, table(dewp), cex=1,pos=3);
boxplot(dewp, horizontal = TRUE, main = "Dew point in Fahrenheit");


text(x=barplot(table(slp),main = "Sea level pressure",ylab = "Frequency"), y = 0, table(slp), cex=1,pos=3);
boxplot(slp, horizontal = TRUE, main = "Sea level pressure");


par(mfrow = c(2,2));
text(x= barplot(table(pcp01),main = "One-hour liquid precipitation",ylab = "Frequency"), y = 0, table(pcp01), cex=1,pos=3);
boxplot(pcp01, horizontal = TRUE, main = "1-hour liquid precipitation");

text(x= barplot(table(pcp06),main = "6-hour liquid precipitation",ylab = "Frequency"), y = 0, table(pcp06), cex=1,pos=3);
boxplot(pcp06, horizontal = TRUE, main = "6-hour liquid precipitation");


par(mfrow = c(1,2));
text(x= barplot(table(pcp24),main = "24-hour liquid precipitation",ylab = "Frequency"), y = 0, table(pcp24), cex=1,pos=3);
boxplot(pcp24, horizontal = TRUE, main = "24-hour liquid precipitation")


par(mfrow = c(2,2));
text(x= barplot(table(start_month),main = "Start Month for Pickups",ylab = "Frequency"), y = 0, table(start_month), cex=1,pos=3);
boxplot(start_month, horizontal = TRUE, main = " Start Month for Pickups ");

text(x= barplot(table(start_day),main = "Start Date for Pickups",ylab = "Frequency"), y = 0, table(start_day), cex=1,pos=3);
boxplot(start_day, horizontal = TRUE, main = " Start Date for Pickups ");


par(mfrow = c(2,2));
text(x= barplot(table(start_hour),main = "Start Hour for Pickups",ylab = "Frequency"), y = 0, table(start_hour), cex=1,pos=3);
boxplot(start_hour, horizontal = TRUE, main = " Start Hour for Pickups ");

text(x= barplot(table(wday),main = " Weekdays ",ylab = "Frequency"), y = 0, table(wday), cex=1,pos=3);

```

## Observations: 
## - 


```{r}
par(mfrow = c(2,2));
hist(pickups, main = "Pickups", xlab = "", col = "dark grey");
boxplot(pickups, horizontal = TRUE, main = "pickups");


hist(spd, main = "Wind speed in miles/hour", xlab = "", col = "dark grey");
boxplot(spd, horizontal = TRUE, main = " Wind speed in miles/hour");
text(fivenum(spd), labels =fivenum(spd), y=1.25);
rug(spd);


par(mfrow = c(2,2));
hist(vsb, main = " Visibility in Miles to nearest tenth", xlab = "",  col = "dark grey");
boxplot(vsb, horizontal = TRUE, main = " Visibility in Miles to nearest tenth");
text(fivenum(vsb), labels =fivenum(spd), y=1.25);
rug(spd);

hist(temp, main = " Temperature in Fahrenheit", xlab = "", col = "dark grey");
boxplot(temp, horizontal = TRUE, main = " Temperature in Fahrenheit");
text(fivenum(spd), labels =fivenum(temp), y=1.25);
rug(temp)


par(mfrow = c(2,2));
hist(temp, main = "Temperature in Fahrenheit", xlab = "", col = "dark grey");
boxplot(temp, horizontal = TRUE, main = " Temperature in Fahrenheit ");
text(fivenum(spd), labels =fivenum(spd), y=1.25);
rug(temp);


hist(dewp, main = " Dew point in Fahrenheit", xlab = "",  col = "dark grey");
boxplot(dewp, horizontal = TRUE, main = " Dew point in Fahrenheit");
text(fivenum(vsb), labels =fivenum(spd), y=1.25);
rug(dewp)


par(mfrow = c(2,2));
hist(slp, main = " Sea level pressure", xlab = "", col = "dark grey");
boxplot(slp, horizontal = TRUE, main = " Temperature in Fahrenheit ");
text(fivenum(slp), labels =fivenum(spd), y=1.25);
rug(slp);

hist(pcp01, main = " 1-hour liquid precipitation", xlab = "",  col = "dark grey");
boxplot(pcp01, horizontal = TRUE, main = " 1-hour liquid precipitation ");
text(fivenum(pcp01), labels =fivenum(spd), y=1.25);
rug(pcp01);


par(mfrow = c(2,2));
hist(pcp06, main = " 6-hour liquid precipitation", xlab = "",  col = "dark grey");
boxplot(pcp06, horizontal = TRUE, main = " 6-hour liquid precipitation ");
text(fivenum(pcp06), labels =fivenum(spd), y=1.25);
rug(pcp06);

hist(pcp24, main = " 24-hour liquid precipitation", xlab = "",  col = "dark grey");
boxplot(pcp24, horizontal = TRUE, main = " 24-hour liquid precipitation ");
text(fivenum(pcp06), labels =fivenum(spd), y=1.25);
rug(pcp24);


par(mfrow = c(2,2));
hist(sd, main = " Snow depth in inches", xlab = "",  col = "dark grey");
boxplot(sd, horizontal = TRUE, main = " Snow depth in inches");
text(fivenum(sd), labels =fivenum(spd), y=1.25);
rug(sd)


# ****************************** Alternative Univariate Analysis **********************************

hist(Uber$vsb, main= "Visibility")
# Almost clear weather 

hist(Uber$temp, main="Temperature")
boxplot(temp,horizontal = TRUE)
# Two peaks culd be seen around 35 and second around 60 in the histogram
# This histogram is bi-modal because it is like 2 hist joint together

## The density plot is gotten by drawing the tops of a histogram. PLOT DENSITY

plot(density(Uber$dewp), main="Dew point variations")
# Distribution looks similar to that of temperature (bi-modal)
# Rememberr, all of these steps are just to look at how the data is behaving, information hunt!

plot(density(Uber$slp), main="Sea level pressure")
# Approximately normaldistribution
# Sea level, dew point(humidity), speed all would be related to temperatures

plot(density(Uber$sd), main="Snow depth")
# No snow for majority of times since the other peaks are way too low and pick is at point 0
# Thus, we think snow depth doesn't affect driving and peakups
# If the number of snow depth was high significantly, it could affect peakups

# ***************************************** END *************************************************************
```

# Bivariate Analysis  - These plots are not very informative
```{r}

# We plot numeric vs categorical variables
boxplot(pickups~borough,horizontal=T,main="Pickups by Boroughs")
levels(borough)
```


## Bivariate - Analysis COntinues 
```{r}
## Check how many holidays in each of the six months in which the data was collected
unique(Uber[which(Uber$hday=="Y"),c("start_day","start_month")])

## Check the number of trips in each month and whether or not there is a holiday in that month 
table(hday,start_month)

plot(aggregate(pickups~hday,data=Uber, mean), type="b")
plot(aggregate(pickups~start_month,data=Uber,mean),type="b")


## This tells us which boroughs get the max number of pickups , in order of preference
## Here, we print the total number of pickups for all the 6 boroughs in NY City

library(ggplot2)
ggplot(Uber,aes(pickups))+
  geom_histogram()+
  facet_wrap(~borough,ncol=3)

## To be very precise using numbers, which boroughs get more pickups, we use the following code 
library(dplyr)
Uber%>%group_by(borough)%>%summarise(Total.Pickups=sum(pickups))%>%arrange(desc(Total.Pickups))


# Check for outliers in variables
boxplot(Uber[,c(4:7)],horizontal = TRUE)
boxplot(Uber[,c(9:12)],horizontal = TRUE)


# *******************************  None of weather variables seem to affect pickups ******************************

# Pickup vs wind speed
plot(spd, pickups, xlab= "speed", ylab="pickup", main ="pickup vs speed")
abline(lm(pickups~spd),col=c("Red"))

# Pick up vs vsb 
plot(vsb, pickups, xlab= "visibility", ylab="pickup", main ="pickup vs visibility")
abline(lm(pickups~vsb),col=c("Red"))

## Precipitation vs pickups
plot(pcp01, pickups, xlab= "pcp01", ylab="pickup", main ="pcp01 vs pickups")
abline(lm(pickups~pcp01),col=c("Red"))

### Temperature vs Pickups
plot(temp, pickups, xlab= "temp", ylab="pickup", main ="temperature vs pickup")
abline(lm(pickups~temp),col=c("Red"))


## Check the other weather factors

# *********************************************** END *********************************************************************


# ******************* Now we see if time based variables will affect the pickups or not ***************************

### Check Pickups in Start_month to see the behavior
plot(aggregate(pickups~start_month,data=Uber, sum), type="b")

## Observation: we see an increase in monthly Pickups. 


## lets see daily bookings of taxi in a month
plot(aggregate(pickups~start_day,data=Uber, sum), type="b", main ="Sum of bookings for each day")

## Observation: Some insights are revealed by the peaks of the graph, pickups were high at those places. 


## Here we exclude the month of february for Pickups and see what graph gives us
Uber %>% 
  filter(.,start_month !=2) %>% 
  ggplot(aes(x=start_day, y=pickups))+geom_bar(stat='identity')


## Pickup on an hourly basis
plot(aggregate(pickups~start_hour,data=Uber, sum), type="b")
## Pickups are very high around the 19th-20th hour and then decrease steadily till 5am and we know why. 


## Pickups based on days of the week
ggplot(aes(x = reorder(wday, pickups), y = pickups), data = Uber) +
  geom_bar(aes(fill=pickups), width=0.5, stat = "identity") + coord_flip()

### Saturday has the highest Pickups in a week followed by Friday and so on..


class(wday)
### Get aggregate pickups for all boroughs of NY City
Uber$wday = as.factor(Uber$wday)
plot(aggregate(pickups~wday,data=Uber, sum), type="b")

## low bookings on monday which rises till saturday and then drops off slightly on sunday


```
## Observation: 
# - There are 2 holidays in January, February, and May. There is 1 holiday in June and no holiday in March and April .
# - There are no holidays in the month's of March and April and hence no trips on those days. The highest number of trips
# - are recorded in March.

# Multivariate Analysis  (NOT NEEDED. It yields redundant information)
```{r}
### We check the bookings per hour for the different boroughs
# geom_jitter gives us the color of the plot
# Unknown found on the bottom, wero, because it is made up of NA's
ggplot(Uber, aes(start_hour, pickups)) +
geom_jitter(alpha = 0.3, aes(colour = borough)) 


## In this case, we have a pattern while immediately above, no!
# geom_smooth gives us the pattern in the diagram
ggplot(Uber, aes(start_hour, pickups)) +
  geom_jitter(alpha = 0.3, aes(colour = borough)) +
geom_smooth(aes(color = borough))

ggplot(Uber, aes(start_hour, borough)) +
  geom_jitter( alpha = 0.4, aes(color = pcp01 > 0)) +
  geom_smooth(aes(color = pcp01 > 0))
```

